---
title: "Covid91 Vaccine Study Final 2023"
author: "Gracie Telles"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

# Males

Does the Covid921 exhibit any differences between male populations?

$H_0$ The Covid921 vaccine does not exhibit a significant difference in effectiveness between male populations.

$H_A$ There is a significant difference in effectiveness of the Covid921 vaccine between male populations.

```{r}
library(dplyr)
Males = filter (FinalData, FinalData$sex == "M")
```


## Methods

Covid921 vaccine administration is an explanatory variable, and Covid921 effectiveness is the response variable. Males is a categorical variable, and Covid 921 vaccine is a categorical variable, therefor we use the CatCat helper file.

## Results

## Graphical Results

```{r}
barchartGC(~infected + treatment,data=Males)
barchartGC(~infected + treatment,data=Males, type="percent")

```

According the the bar graph the Males who got the placebo had a higher percentage of getting Covid 19, as opposed to those who received the vaccine. The Men who received the vaccine  had a significant reduction in catching Covid. 

## Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=Males)
rowPerc(table1)

```

The table verifies that more than double of the people that were infected by Covid 19 were given the placebo. Of those that did not get sick, placebo and vaccine treatment was almost equal.

## Inferential

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```


In Males, those who did not get sick with Covid, it was about 50/50 for who received the vaccine and who received a placebo. The more telling data indicates that of those who did get sick, those that were given the placebo were twice as likely as those that received the vaccine. 

# Females

Does the Covid921 exhibit any differences between female populations?

$H_0$ The Covid921 vaccine does not exhibit a significant difference in effectiveness between female populations.

$H_A$ There is a significant difference in effectiveness of the Covid921 vaccine between female populations.

```{r}
library(dplyr)
Females = filter (FinalData, FinalData$sex == "F")
```


## Methods

Covid921 vaccine administration is an explanatory variable, and Covid921 effectiveness is the response variable. Females is a categorical variable, and Covid 921 vaccine is a categorical variable, therefor we use the CatCat helper file.

## Results

## Graphical Results

```{r}
barchartGC(~infected + treatment,data=Females)
barchartGC(~infected + treatment,data=Females, type="percent")

```

According the the bar graph the Females who got the placebo had a higher percentage of getting Covid 19, as opposed to those who received the vaccine. The Men who received the vaccine  had a significant reduction in catching Covid. 

## Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=Females)
rowPerc(table1)

```

## Inferential

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```


# LGBTQ

Does the Covid921 exhibit any differences between LGBTQ populations?

$H_0$ The Covid921 vaccine does not exhibit a significant difference in effectiveness between LGBTQ populations.

$H_A$ There is a significant difference in effectiveness of the Covid921 vaccine between LGBTQ populations.

```{r}
library(dplyr)
LGBTQ = filter (FinalData, FinalData$LGBTQ == "gay")
```

## Methods

Covid921 vaccine administration is an explanatory variable, and Covid921 effectiveness is the response variable. LGBTQ is a categorical variable, and Covid 921 vaccine is a categorical variable, therefor we use the CatCat helper file.

## Results 

## Graphical Results

```{r}
barchartGC(~infected + treatment,data=LGBTQ)
barchartGC(~infected + treatment,data=LGBTQ, type="percent")

```

According the the bar graph the LGBTQ people who got the placebo had a lower percentage of getting Covid 19, as opposed to those who received the vaccine. The LGBTQ who received the vaccine  had a significant reduction in catching Covid. 

## Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=LGBTQ)
rowPerc(table1)

```

## Inferential

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```



# Drug Users

Does the Covid921 exhibit any differences between drug users populations?

$H_0$ The Covid921 vaccine does not exhibit a significant difference in effectiveness between drug user populations.

$H_A$ There is a significant difference in effectiveness of the Covid921 vaccine between drug user populations.

```{r}
library(dplyr)
Druggies = filter (FinalData, FinalData$DrugUser == "yes")
```


## Methods

Covid921 vaccine administration is an explanatory variable, and Covid921 effectiveness is the response variable. Drug users is a categorical variable, and Covid 921 vaccine is a categorical variable, therefor we use the CatCat helper file.

## Results

## Graphical Results

```{r}
barchartGC(~infected + treatment,data=Druggies)
barchartGC(~infected + treatment,data=Druggies, type="percent")

```

## Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=Druggies)
rowPerc(table1)

```

## Inferential

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```


# Overall Results and Conclusions