---
title: "Covid921 Vaccine Study Final 2023"
author: "Gracie Telles"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

# Males

Does the Covid921 exhibit any differences between male populations? Is the vaccine effective against the virus?

$H_0$ The Covid921 vaccine does not exhibit a significant difference in effectiveness between male populations.

$H_A$ There is a significant difference in effectiveness of the Covid921 vaccine between male populations.

```{r}
library(dplyr)
Males = filter (FinalData, FinalData$sex == "M")
```


## Methods

Covid921 vaccine administration is an explanatory variable, and Covid921 effectiveness is the response variable. Males is a categorical variable, and Covid 921 vaccine is a categorical variable, therefor we use the CatCat helper file.

## Results

## Graphical Results

```{r}
barchartGC(~infected + treatment,data=Males)
barchartGC(~infected + treatment,data=Males, type="percent")

```

According the the bar graph the Males who got the placebo had a higher percentage of getting Covid 19, as opposed to those who received the vaccine. The Men who received the vaccine had a significant reduction in catching Covid. 

## Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=Males)
rowPerc(table1)

```

The table verifies that more than double of the people that were infected by Covid 19 were given the placebo. Of those that did not get sick, placebo and vaccine treatment were almost equal.

## Inferential

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```


The data presented indicates that of those who did get sick, they were twice as likely to have had the placebo as opposed to the vaccine. The P Value for the vaccine is 0.000000000000198, which is staggeringly lower than 0.050. The odds ration is 2.24, and the bar graphs and table provide the same telling data: there is a significant difference in the effectiveness of the vaccine in male populations, therefor we reject the null hypothesis. The effectiveness of the vaccine in male populations is positive, and I would suggest using Covid921 as a vaccine for that subgroup.

# Females

Does the Covid921 exhibit any differences between female populations? Is the vaccine effective against the virus?

$H_0$ The Covid921 vaccine does not exhibit a significant difference in effectiveness between female populations.

$H_A$ There is a significant difference in effectiveness of the Covid921 vaccine between female populations.

```{r}
library(dplyr)
Females = filter (FinalData, FinalData$sex == "F")
```


## Methods

Covid921 vaccine administration is an explanatory variable, and Covid921 effectiveness is the response variable. Females is a categorical variable, and Covid 921 vaccine is a categorical variable, therefor we use the CatCat helper file.

## Results

## Graphical Results

```{r}
barchartGC(~infected + treatment,data=Females)
barchartGC(~infected + treatment,data=Females, type="percent")

```

According the the bar graph the Females who got the placebo had a higher percentage of getting Covid 19, as opposed to those who received the vaccine. The Females who received the vaccine had a significant reduction in catching Covid. 

## Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=Females)
rowPerc(table1)

```

The table verifies that 60% of the people that were infected by Covid 19 were given the placebo, as opposed to the 40% that were vaccinated. Of those that did not get sick, placebo and vaccine treatment were almost equal.

## Inferential

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The data presented indicates that of those that did get sick, 60% of them had received the placebo. The P Value for the vaccine is 0.0002732, which is lower than 0.050. The odds ration is 1.49, and the bar graphs and table provide the same telling data: there is a significant difference in the effectiveness of the vaccine in Female populations, therefor we reject the null hypothesis. The effectiveness of the vaccine in Female populations is positive, and I would suggest using Covid921 as a vaccine for that subgroup.

# LGBTQ

Does the Covid921 exhibit any differences between LGBTQ populations? Is the vaccine effective against the virus?

$H_0$ The Covid921 vaccine does not exhibit a significant difference in effectiveness between LGBTQ populations.

$H_A$ There is a significant difference in effectiveness of the Covid921 vaccine between LGBTQ populations.

```{r}
library(dplyr)
LGBTQ = filter (FinalData, FinalData$LGBTQ == "gay")
```

## Methods

Covid921 vaccine administration is an explanatory variable, and Covid921 effectiveness is the response variable. LGBTQ is a categorical variable, and Covid 921 vaccine is a categorical variable, therefor we use the CatCat helper file.

## Results 

## Graphical Results

```{r}
barchartGC(~infected + treatment,data=LGBTQ)
barchartGC(~infected + treatment,data=LGBTQ, type="percent")

```

According the the bar graph the overwhelming majority of LGBTQ people who were infected with the virus received the vaccine. 

## Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=LGBTQ)
rowPerc(table1)

```

According to the table, about 84% of vaccinated LGBTQ people became infected with the virus. Of those that did not get sick, placebo and vaccine treatment were almost equal.

## Inferential

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The data presented indicates that of those LGBTQ people that did get sick, 84% of them had been vaccinated. The P Value for the vaccine is 0.00000006, which is lower than 0.050. The odds ration is 0.18, and the bar graphs and table provide the same telling data: there is a significant difference in the effectiveness of the vaccine in LGBTQ populations, therefor we reject the null hypothesis. The effectiveness of the vaccine in LGBTQ populations is poor, and because of its negative ability to keep them healthy, I would not suggest using Covid921 as a vaccine for that subgroup.

# Drug Users

Does the Covid921 exhibit any differences between drug user populations? Is the vaccine effective against the virus?

$H_0$ The Covid921 vaccine does not exhibit a significant difference in effectiveness between drug user populations.

$H_A$ There is a significant difference in effectiveness of the Covid921 vaccine between drug user populations.

```{r}
library(dplyr)
Druggies = filter (FinalData, FinalData$DrugUser == "yes")
```


## Methods

Covid921 vaccine administration is an explanatory variable, and Covid921 effectiveness is the response variable. Drug users is a categorical variable, and Covid 921 vaccine is a categorical variable, therefor we use the CatCat helper file.

## Results

## Graphical Results

```{r}
barchartGC(~infected + treatment,data=Druggies)
barchartGC(~infected + treatment,data=Druggies, type="percent")

```

According the the bar graph the overwhelming majority of Drfug Using people who were infected with the virus received the vaccine. 

## Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=Druggies)
rowPerc(table1)

```

According to the table, about 88% of vaccinated Drug Users became infected with the virus. Of those that did not get sick, placebo and vaccine treatment were almost equal.

## Inferential

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The data presented indicates that of those Drug Users that did get sick, 88% of them had been vaccinated. The P Value for the vaccine is 0.000000002, which is lower than 0.050. The odds ration is 0.12, and the bar graphs and table provide the same telling data: there is a significant difference in the effectiveness of the vaccine in the Drug User population, therefor we reject the null hypothesis. The effectiveness of the vaccine in Drug Users is poor, and because of its negative ability to keep them healthy, I would not suggest using Covid921 as a vaccine for that subgroup.

# Overall Results and Conclusions

The original question was if Covid 921 exhibited any differences in the Male, Female, LGBTQ and Drug User populations, and if it was an effective vaccine. The data in the previous graphs, tables and Chisq and Fishers test all have the same conclusion: yes, there IS a difference exhibited between those populations that received the vaccine versus those that received teh placebo. Therefor we reject the NULL hypothesis. However, the effectiveness of the vaccine within the subgroups was varied. With both Male and Female populations, the vaccine was overwhelmingly successful, and should be used for the further prevention of Covid 19. Unfotunately, the results are the exact opposite for the LGBTQ and Drug Users tested. In those cases, the effectiveness of the vaccine was overwhelmingly poor, and should definitely not be used in the future as a means for preventing the infection of those subgroups. 