---
title: "Covid91 Vaccine Study Final 2023"
author: "Gracie Telles"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

# Males

$H_0$ The Covid921 vaccine does not exhibit a significant difference in effectiveness between male populations.

$H_A$ There is a significant difference in effectiveness of the Covid921 vaccine between male populations.

```{r}
library(dplyr)
Males = filter (FinalData, FinalData$sex == "M")
```


## Methods

Covid921 vaccine administration is an explanatory variable, and Covid921 effectiveness is the response variable. Males is a categorical variable, and Covid 921 vaccine is a categorical variable, therefor we use the CatCat helper file.

## Descriptive Results



## Graphical Results

```{r}
barchartGC(~infected + treatment,data=Males)
barchartGC(~infected + treatment,data=Males, type="percent")

```

According the the bar graph the Males who got the placebo had a higher percentage of getting Covid 19, as opposed to those who received the vaccine. The Men who received the vaccine  had a significant reduction in catching Covid. 

## Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=Males)
rowPerc(table1)

```

## Inferential


# Females

$H_0$ The Covid921 vaccine does not exhibit a significant difference in effectiveness between female populations.

$H_A$ There is a significant difference in effectiveness of the Covid921 vaccine between female populations.

```{r}
library(dplyr)
Females = filter (FinalData, FinalData$sex == "F")
```


## Methods

Covid921 vaccine administration is an explanatory variable, and Covid921 effectiveness is the response variable. Females is a categorical variable, and Covid 921 vaccine is a categorical variable, therefor we use the CatCat helper file.

## Descriptive Results

## Graphical Results

```{r}
barchartGC(~infected + treatment,data=Females)
barchartGC(~infected + treatment,data=Females, type="percent")

```

According the the bar graph the Females who got the placebo had a higher percentage of getting Covid 19, as opposed to those who received the vaccine. The Men who received the vaccine  had a significant reduction in catching Covid. 

## Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=Females)
rowPerc(table1)

```

## Inferential


# LGBTQ

$H_0$ The Covid921 vaccine does not exhibit a significant difference in effectiveness between LGBTQ populations.

$H_A$ There is a significant difference in effectiveness of the Covid921 vaccine between LGBTQ populations.

```{r}
library(dplyr)
LGBTQ = filter (FinalData, FinalData$LGBTQ == "gay")
```

## Methods

Covid921 vaccine administration is an explanatory variable, and Covid921 effectiveness is the response variable. LGBTQ is a categorical variable, and Covid 921 vaccine is a categorical variable, therefor we use the CatCat helper file.

## Descriptive Results

## Graphical Results

```{r}
barchartGC(~infected + treatment,data=LGBTQ)
barchartGC(~infected + treatment,data=LGBTQ, type="percent")

```

According the the bar graph the LGBTQ people who got the placebo had a lower percentage of getting Covid 19, as opposed to those who received the vaccine. The LGBTQ who received the vaccine  had a significant reduction in catching Covid. 

## Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=LGBTQ)
rowPerc(table1)

```

## Inferential


# Druggies

$H_0$ The Covid921 vaccine does not exhibit a significant difference in effectiveness between druggie populations.

$H_A$ There is a significant difference in effectiveness of the Covid921 vaccine between druggie populations.

```{r}
library(dplyr)
Druggies = filter (FinalData, FinalData$DrugUser == "yes")
```


## Methods

Covid921 vaccine administration is an explanatory variable, and Covid921 effectiveness is the response variable. Drug users is a categorical variable, and Covid 921 vaccine is a categorical variable, therefor we use the CatCat helper file.

## Descriptive Results

## Graphical Results

```{r}
barchartGC(~infected + treatment,data=Druggies)
barchartGC(~infected + treatment,data=Druggies, type="percent")

```

## Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=Druggies)
rowPerc(table1)

```

## Inferential


# Overall Results and Conclusions